---
title: "Report 1 431 project1"
author: "William Li"
format: html
---

**Context of Data**
The data set I will be using captures the global audit results from plastic waste that is collected from volunteers. The waste was then used to link pollution back to specific companies and the materials found to better learn how to combat the pollution 

**What cleaning has been done**

The Janitor function was used to clean the column names and standardize them, mutate was used to convert strings into numeric and add a year column. Country names were extracted and stored in a new column for the 2019 data set. In the 2019 data set duplicate columns were removed also inconsistent column names were renamed. Finally the data sets were combined into one singular data set 

**Research question 1**

What companies are the largest contributors to plastic waste and pollution

**Research question 2**

Does plastic composition vary across countries

```{r}
install.packages("tidytuesdayR")

tuesdata <- tidytuesdayR::tt_load('2021-01-26')
tuesdata <- tidytuesdayR::tt_load(2021, week = 5)

plastics <- tuesdata$plastics


plastics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-01-26/plastics.csv')

```


**Visualization 1**
```{r}
library(tidyverse) 

plastics %>%
  group_by(parent_company) %>%
  summarise(total = sum(grand_total, na.rm = TRUE)) %>%
  slice_max(total, n = 10) %>%
  ggplot(aes(x = reorder(parent_company, total), y = total)) +
  geom_col() +
  coord_flip() +
  labs(title = "Top 10 Companies", x = "Company", y = "Total Plastic")
```


**visualization 2**
```{r}
plastics %>%
  group_by(year) %>%
  summarise(total = sum(grand_total, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = total)) +
  geom_line() +
  geom_point() +
  labs(title = "Total Plastic Pollution Over Time",
       x = "Year",
       y = "Total Plastic")
```

